### 2012-03-30 ms.tex
Initial commit

\documentclass[letterpaper, 11pt]{article}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[left=3cm,top=3cm,right=3cm]{geometry}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\parindent=0cm

\title{Robust Inferences from Simply Parameterised Models}

\author{%
  Brendon~J.~Brewer$^{1}$\thanks{\tt brewer@physics.ucsb.edu}, 
  Geraint~F.~Lewis$^{2}$,
  Ross Fadely\\
  \medskip\\
  $^1$UCSB\\
  $^2$USyd
}

\begin{document}

\date{\today}

\maketitle

\begin{abstract}
We discuss the practical use of simply parameterised models in Bayesian data analysis, particularly in astronomy.
The use of simply parameterised models is problematic when they are fit to large, informative data sets, resulting in unrealistically small uncertainties on inferences. In common parlance, the formal statistical errors may be small, but the systematic errors are often large and in many cases, difficult to quantify at all. In principle, the best solution to this problem is to develop more realistic complex models that accurately describe our prior beliefs. However, this is often infeasible in practice, due to computational limitations and the difficulty of defining realistic prior distributions over high dimensional spaces. As a pragmatic alternative, we show that it is possible to reinterpret the use of simple models in a way that avoids the problem of overconfident
inferences. The key idea is to weaken the connection between the parameters and the data that is assumed in the prior probabilities. This weakening can be implemented straightforwardly using the Nested Sampling algorithm.
\end{abstract}

\section{Background on Bayesian Inference}

\begin{itemize}
\item Define terminology very carefully, present Bayes' rule \\
\item Describe the ``joint'' (parameters $\times$ data space) view of inference, describe how the form of the likelihood function is prior information \citep{caticha} \\
\item Discuss use of MCMC techniques for complex problems \\
\item Discuss the various attempts at robustness/model checking/elicitation and that stuff
\end{itemize}

\begin{equation}
p(\theta | D) \propto p(\theta) p(D|\theta)|_{D=D_{\rm obs}}
\end{equation}
The various terms in this equation have the following names and meanings.

\begin{itemize}
\item $p(\theta|D)$ is the {\it posterior distribution}
for the parameters, and is usually the goal of the calculation. It describes our
knowledge and remaining uncertainty about $\theta$ after taking into account the data. \\

\item $p(\theta)$ is the {\it prior distribution}, a probability distribution
over possible values for $\theta$ that describes our uncertainty about
$\theta$ before taking into account the data. Commonly, this is a broad
probability distribution, describing a large amount of uncertainty. \\

\item $p(D|\theta)$ is a function of
both $\theta$ and $D$ and is referred to as the {\it sampling distributions}. 
The reason for the awkward plural is that $p(D|\theta)$ is a set of probability distributions,
one for each value of $\theta$. The sampling distributions describe our prior information about how the data are connected to the parameters \citep{2008arXiv0808.0012C}. \\

\item $p(D|\theta)|_{D = D_{\rm obs}}$ is called the {\it likelihood function}.
It is obtained from the sampling distributions, with the actual data substituted in,
and is therefore a function of $\theta$ only. It describes the effects of the actual data
set that was observed on our beliefs about $\theta$: values of $\theta$ with a high
likelihood predicted the data well and will have their probability upweighted
in the posterior distribution.
\end{itemize}

\begin{figure*}
\begin{center}
\includegraphics[scale=0.5]{joint_marginal.eps}
\end{center}
\end{figure*}

\section{Advantages and Disadvantages of Simply Parameterised Models}
\begin{itemize}
\item Easy to implement \\
\item Computationally fast to get inferences - therefore can be applied on a large scale, e.g. to large samples of objects. \\
\item Easy to assign non-controversial priors over low dimensional parameter spaces \\
\end{itemize}

\section{Problems with Simply Parameterised Models}
\begin{itemize}
\item They are usually false, i.e. poor models of our actual prior beliefs. \\
\item Conditional on the simple model, you get error bars that are much too small. Small error bars on the parameters may not matter so much (since the parameters don't ``exist'' anyway), but the inferences on quantities that do ``exist'' (e.g. total flux) will still be overconfident. \\
\end{itemize}

\section{Re-Interpreting Simply Parameterised Models}

\begin{itemize}
\item Describe the hypothetical process of acknowledging the complex reality, yet describing (summarising) it by fitting a simple model to the complex one. \\
\item Show some examples of this - complex curves being approximated by lines with least squares, complex density profiles (1-D and 2-D) being fit by simple density profiles by maximising some kind of entropy.
\item Mention/derive that in the special case of fitting a complex density profile with a Gaussian, the variance(s) of the Gaussian gives the actual moment(s) of the complex profile
\item Describe how fitting a simple model can be interpreted as attempting to infer the parameters of the summarising simple model, and that therefore the parameter space is completely legitimate - it's the likelihood function that is now wrong. Data = model predictions + noise breaks down, it should be Data = simple model predictions + complexity perturbations + noise
\item Describe why qualitatively, the likelihood function should now be ``weaker'' than before
\item Describe why the old likelihood function is still (often) a good hill to climb
\end{itemize}

\begin{figure*}
\begin{center}
\includegraphics[scale=0.25]{simple_complex1.eps}
\includegraphics[scale=0.25]{simple_complex2.eps}
\includegraphics[scale=0.25]{simple_complex3.eps}
\end{center}
\end{figure*}

\section{Implementation with Nested Sampling}
\begin{itemize}
\item Describe the concept of tempering/annealing and how it almost does what we want \\
\item But we don't really know what temperature we want at the outset - NS allows us to test all in a single run
\item Introduce ``convergence plots'' and how to read them
\item Subjective temperature selection and why it's not so bad
\end{itemize}

\begin{figure*}
\begin{center}
\includegraphics[scale=0.7]{nested.eps}
\end{center}
\end{figure*}

\section{More Objective Temperature Selection}
\begin{itemize}
\item If we can generate from a prior over complex models, and then fit simple models to each sample, and also generate simulated data from the complex model, we can get an idea of the joint prior over the space of (simple parameters, data)\\
\item Approximate the joint prior over (simple parameters, data) with a tempered version of the old likelihood \\
\item Find the optimal temperature that makes the approximated joint prior as close as possible to the true one.
\end{itemize}

\section{Failure Modes}
\begin{itemize}
\item If the simple model / original likelihood function is SO BAD that optimising it takes you to the wrong place in parameter space
\end{itemize}

\section{Simple Example 1: A rotation curve}

\section{Simple Example 2: An emission line}

\section{Potential Applications}

\begin{itemize}
\item Interferometry: a common goal of interferometry is to measure the angular size of some astronomical object, when it is too small to be resolved by standard telescopes. In the case of stellar interferometry, the surface brightness profile of a star is very well modelled by a circular disc with limb darkening, so this model should be used. However, with other objects, such as the dust torii in active galactic nuclei \citep[e.g.][]{2011A&A...536A..78K}, the spatial profile is not well known. If a characteristic size, or total flux or something, is the goal of the study, then the framework of this paper is applicable.\\

\item Gravitational lensing: Modelling of gravitational lenses typically aims to infer the source surface brightness profile along with the projected mass density profile of the deflector. While sophisticated many-parameter models are feasible and are used routinely now \citep{2006MNRAS.371..983S, 2007ApJ...666..726B, 2011MNRAS.412.2521B, 2006ApJ...637..608B, 2012Natur.481..341V}, they are slow, and some are not particularly realistic anyway (rant about ``regularisation'' here). With the explosion in the number of known gravitational lens systems (ref?), population-based studies will need something faster. I have already used this approach \citep{2012arXiv1201.1677B}.

\item Galaxy Structure and Morphology: Studies of populations of galaxies are often carried out by fitting simple models such as the S{\'e}rsic \citep{Sersic1968} profile to individual galaxies. It has been demonstrated that taking into account the full uncertainties on these parameters can yield different results as compared to simply studying the distribution of the best-fit parameters \citep[e.g.][]{2011MNRAS.414.1625Y}. However, since galaxies often are not well modelled by S{\'e}rsic profiles (due to their real-world complexity), the error bars on the parameters can be severely underestimated, which is almost like just using the best-fit parameters! \\

\item Studying spectra: Measuring equivalent widths, velocity dispersions and so on from emission/absorption lines. These commonly assume a specific functional form for the spectral line profile (e.g. Gaussian), which may not be applicable.

\item Any situation with correlated noise (but a believed model), we can use this method as well (e.g. \citep{2008ApJ...686..851R}).

\item Ideas? There are surely a large number of possibilities
\end{itemize}

\section*{Acknowledgements}
Tommaso Treu, Anna Pancoast, Phil Marshall, Matthew Auger, Greg Dobler, Sebastian Hoenig

\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Barnab{\`e} 
\& Koopmans}{2007}]{2007ApJ...666..726B} Barnab{\`e} M., Koopmans L.~V.~E., 2007, ApJ, 666, 726

\bibitem[\protect\citeauthoryear{Brewer 
\& Lewis}{2006}]{2006ApJ...637..608B} Brewer B.~J., Lewis G.~F., 2006, ApJ, 637, 608

\bibitem[\protect\citeauthoryear{Brewer et al.}{2011}]{2011MNRAS.412.2521B} 
Brewer B.~J., Lewis G.~F., Belokurov V., Irwin M.~J., Bridges T.~J., Evans 
N.~W., 2011, MNRAS, 412, 2521

\bibitem[\protect\citeauthoryear{Brewer et al.}{2012}]{2012arXiv1201.1677B} 
Brewer B.~J., et al., 2012, arXiv, arXiv:1201.1677

\bibitem[\protect\citeauthoryear{Caticha}{2008}]{caticha} 
Caticha A., 2008, arXiv, arXiv:0808.0012

\bibitem[\protect\citeauthoryear{Jaynes 
\& Bretthorst}{2003}]{2003prth.book.....J} Jaynes E.~T., Bretthorst G.~L., 2003, prth.book,

\bibitem[\protect\citeauthoryear{Kishimoto et 
al.}{2011}]{2011A&A...536A..78K} Kishimoto M., H{\"o}nig S.~F., Antonucci R., Millour F., Tristram K.~R.~W., Weigelt G., 2011, A\&A, 536, A78

\bibitem[\protect\citeauthoryear{Riechers et 
al.}{2008}]{2008ApJ...686..851R} Riechers D.~A., Walter F., Brewer B.~J., 
Carilli C.~L., Lewis G.~F., Bertoldi F., Cox P., 2008, ApJ, 686, 851

\bibitem[{{S{\'e}rsic}(1968)}]{Sersic1968}
{S{\'e}rsic}, J.~L. 1968, {Atlas de galaxias australes} (Cordoba, Argentina:
  Observatorio Astronomico)

\bibitem[\protect\citeauthoryear{Sivia \& Skilling}{2006}]{sivia} Sivia, 
D.~ S., Skilling, J., 2006, Data Analysis: A Bayesian Tutorial, 2nd 
Edition, Oxford University Press

\bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, 
J., 2006, ``Nested Sampling for General Bayesian Computation'', Bayesian 
Analysis 4, pp. 833-860

\bibitem[\protect\citeauthoryear{Suyu et al.}{2006}]{2006MNRAS.371..983S} 
Suyu S.~H., Marshall P.~J., Hobson M.~P., Blandford R.~D., 2006, MNRAS, 
371, 983

\bibitem[\protect\citeauthoryear{Vegetti et 
al.}{2012}]{2012Natur.481..341V} Vegetti S., Lagattuta D.~J., McKean J.~P., 
Auger M.~W., Fassnacht C.~D., Koopmans L.~V.~E., 2012, Natur, 481, 341

\bibitem[\protect\citeauthoryear{Yoon, Weinberg, 
\& Katz}{2011}]{2011MNRAS.414.1625Y} Yoon I., Weinberg M.~D., Katz N., 2011, MNRAS, 414, 1625 
\end{thebibliography}

\end{document}