### 2013-12-28 ms2.tex
New document

\documentclass[a4paper, 11pt]{article}
\usepackage{authblk}	% For nice author-affil lists
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citet and citep
\usepackage{amsmath}	% for \iint
\usepackage{bbm}	% for blackboard bold numbers
\usepackage[left=3cm,top=3cm,right=3cm]{geometry}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\parindent=0cm
\newcommand{\yy}{\mathbf{y}}
\newcommand{\dnest}{\texttt{DNest}}

\title{Inferring Macroscopic Properties}

\author[1]{Brendon J. Brewer}
\affil[1]{Department of Statistics, University of Auckland, Private Bag 92019, Auckland, New Zealand}

\begin{document}

\date{\today}

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}
A hypothesis space indexed by parameters
$\mathbf{x} = \{x_1, x_2, ..., x_n\}$. A macroscopic quantity
is defined by a scalar function $f(\mathbf{x})$.

\section*{Acknowledgements}
Tommaso Treu, Anna Pancoast, Phil Marshall, Matthew Auger, Greg Dobler, Sebastian Hoenig, Douglas Applegate

\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Barnab{\`e} 
\& Koopmans}{2007}]{2007ApJ...666..726B} Barnab{\`e} M., Koopmans L.~V.~E., 2007, ApJ, 666, 726

\bibitem[\protect\citeauthoryear{Brewer 
\& Lewis}{2006}]{2006ApJ...637..608B} Brewer B.~J., Lewis G.~F., 2006, ApJ, 637, 608

\bibitem[\protect\citeauthoryear{Brewer et al.}{2011}]{2011MNRAS.412.2521B} 
Brewer B.~J., Lewis G.~F., Belokurov V., Irwin M.~J., Bridges T.~J., Evans 
N.~W., 2011, MNRAS, 412, 2521

\bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay, 
\& Cs{\'a}nyi}{2011}]{dnest} {\bf Brewer B.~J.}, P{\'a}rtay L.~B., Cs{\'a}nyi G., 2011, Diffusive Nested Sampling, 
Statistics and Computing, 21, 4, 649-656.

\bibitem[\protect\citeauthoryear{Brewer et al.}{2012}]{2012arXiv1201.1677B} 
Brewer B.~J., et al., 2012, arXiv, arXiv:1201.1677

\bibitem[\protect\citeauthoryear{Caticha}{2008}]{caticha} 
Caticha A., 2008, arXiv, arXiv:0808.0012

\bibitem[Gregory(2005)]{gregory} Gregory, P.~C., 2005, 
Bayesian Logical Data Analysis for the Physical Sciences. Cambridge 
University Press.\

\bibitem[\protect\citeauthoryear{Jaynes 
\& Bretthorst}{2003}]{2003prth.book.....J} Jaynes E.~T., Bretthorst G.~L., 2003, prth.book,

\bibitem[\protect\citeauthoryear{Kishimoto et 
al.}{2011}]{2011A&A...536A..78K} Kishimoto M., H{\"o}nig S.~F., Antonucci R., Millour F., Tristram K.~R.~W., Weigelt G., 2011, A\&A, 536, A78

\bibitem[\protect\citeauthoryear{Riechers et 
al.}{2008}]{2008ApJ...686..851R} Riechers D.~A., Walter F., Brewer B.~J., 
Carilli C.~L., Lewis G.~F., Bertoldi F., Cox P., 2008, ApJ, 686, 851

\bibitem[{{S{\'e}rsic}(1968)}]{Sersic1968}
{S{\'e}rsic}, J.~L. 1968, {Atlas de galaxias australes} (Cordoba, Argentina:
  Observatorio Astronomico)

\bibitem[\protect\citeauthoryear{Sivia \& Skilling}{2006}]{sivia} Sivia, 
D.~ S., Skilling, J., 2006, Data Analysis: A Bayesian Tutorial, 2nd 
Edition, Oxford University Press

\bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, 
J., 2006, ``Nested Sampling for General Bayesian Computation'', Bayesian 
Analysis 4, pp. 833-860

\bibitem[\protect\citeauthoryear{Suyu et al.}{2006}]{2006MNRAS.371..983S} 
Suyu S.~H., Marshall P.~J., Hobson M.~P., Blandford R.~D., 2006, MNRAS, 
371, 983

\bibitem[Richards et al.(2011)]{2011arXiv1105.6344R} Richards, J.~W., Lee, 
A.~B., Schafer, C.~M., Freeman, P.~E.\ 2011.\ Prototype selection for 
parameter estimation in complex models.\ ArXiv e-prints arXiv:1105.6344.

\bibitem[\protect\citeauthoryear{Vegetti et 
al.}{2012}]{2012Natur.481..341V} Vegetti S., Lagattuta D.~J., McKean J.~P., 
Auger M.~W., Fassnacht C.~D., Koopmans L.~V.~E., 2012, Natur, 481, 341

\bibitem[\protect\citeauthoryear{Yoon, Weinberg, 
\& Katz}{2011}]{2011MNRAS.414.1625Y} Yoon I., Weinberg M.~D., Katz N., 2011, MNRAS, 414, 1625 
\end{thebibliography}

\end{document}